\thetitle

### 2020-11-21 writing/discussion.tex
Add outline items to discussion outline
Add draft of discussion

\begin{itemize}
  \item Interesting discoveries...?

  \item Comparison of existing GT systems represented in our study.  
    Natural is here, transient is here, erasure is here, and they
        compare as...

      \item Threats to validity:
        \begin{itemize}
          \item Racket stack traces
          \item Mutant sampling
          \item Blame trails do not necessarily capture all aspects of error reporting
          \item Transient blame adaptations
          \item More?

        \end{itemize}
\end{itemize}

@@

\subsection{Threat: Is our Transient Correct?}
%% purpose: talk about transient-blame implementation .... what did we do,
%%  how did we test, what threats remain?
%%
%% the benchmark timeouts look like a threat, but this section shows that
%%  we've looked carefully at those

Surprisingly, Transient blames an empty set of boundaries in XXXX cases.
Such cases should never occur, at least in theory, because an empty set
 means the value has never crossed a boundary---if the value is indeed defined
 in the current module, then we have typed code blaming itself for a typed
 value; thus, the type checker must be unsound.
After careful investigation of these empty blame cases, we found no soundness
 bugs in Typed Racket.
Instead, we found scenarios in which Transient lost track of the proper
 boundaries.
These scenarios suggest ways to improve the blame algorithm from~\citet{vss-popl-2017}:
\begin{itemize}
  \item
    Entries in a blame map must point to several parent entries.
    For example, if the function \texttt{f} receives bad input in the call
    \texttt{(filter f xs)}, then blame should point to both \texttt{filter}
    and the \texttt{xs} list.
  \item
    The construction of blame map entries must be guided by type-like specifications
     instead of relying on syntax.
    Aliasing the built-in \texttt{filter} function should not change the shape
     of the blame map.
  \item
    The initial blame map must reflect the initial type environment.
    Typed Racket trusts that untyped core-library functions, such as \texttt{filter},
     behave correctly, but does not list the assumption in the blame map.
\end{itemize}
\noindent{}Despite these known limitations, we conjecture that these improvements
 to Transient blame-tracking will not affect our overall conclusions.

>>>>

%% \begin{itemize}
%% \item Why does the data look like this?
%%   - why tr fails
%%     - bad runtime errors
%%     - TR bug
%%   - why transient fails
%%     - bad runtime errors
%%     - timeouts/ooms: give context wrt fully untyped/typed versions
%%     - empty blame: trim Ben's prose

%%   \item Any kind of dynamic checking has a large positive impact, whether blame or stacktraces are used.
%%     \begin{itemize}
%%     \item Both in terms of usefulness and programmer effort
%%     \item Erasure extremely often fails to provide any helpful information.
%%     \item That said, there are no long trails for Erasure.
%%     \end{itemize}

%%   \item Blame helps significantly over just using stacktraces for both Natural and Transient.
%%     \begin{itemize}
%%     \item Natural's blame improves over exceptions by nearly 20\%
%%     \item Transient's blame improves over exceptions half as much
%%     \end{itemize}

%%   \item Natural clearly offers the most utility, but Transient is not too far behind.
%%     \begin{itemize}
%%     \item Natural clearly improves over all other modes
%%     \item Transient improves significantly over all other modes except Natural.
%%     \item Keep in mind that our selection of scenarios is biased toward Erasure & Transient.
%%     \end{itemize}

%%   \item Discussion of how confidence / margin of error should be interpreted wrt these results?

%%    \item Threats to validity:
%%         \begin{itemize}
%%           \item Racket stack traces
%%           \item Mutant sampling
%%           \item Blame trails do not necessarily capture all aspects of error reporting
%%           \item Transient blame adaptations
%%           \item More?

%%    \end{itemize}
%% \end{itemize}

\subsection{Why does the data look like this?}

The figures in the last section raise a key question: why does blame fail in our dataset?
We analyze the reasons for failure separately for each mode of the rational programmer.

Breaking down the reasons for failure for all scenarios in which Natural is unsuccessful reveals two causes.
First, most of the failing scenarios begin with a runtime exception, which the rational programmer can use to type components until blame becomes available.
In these failing cases (there are 1708), however, blame never becomes available; the exception information points to the wrong component and the trail ends in failure.
A second, much smaller set (40 scenarios) do produce a runtime type error blaming some component, but the blame identifies a non-buggy typed component.
After careful inspection of these scenarios, we are confident that they reveal a bug in the implementation of Typed Racket.

Transient's blame failures break down in a similar way.
Like Natural, most failures occur with unhelpful exception information without ever getting blame (1851 scenarios).
However, Transient also has a substantial number of failures that occur because the programs hit the resource limits (timed out or ran out of memory) and had to be killed (some 770 scenarios).
Additionally, there are nearly a thousand cases where Transient's checks catch a problem, but the blame set reported is empty.
We discuss each of these last two failure categories in detail in~\ref{sec:threat:transient}.

\subsection{High level conclusions}

The analysis in the last section indicates a number of interesting conclusions about blame in the gradual typing setting.
First, figure~\ref{fig:avo-matrix} makes it clear that any kind of dynamic checking has a large positive impact over Erasure, and this is true regardless of whether the rational programmer uses blame or stacktraces.
Indeed, the bottom row of figure~\ref{fig:avo-matrix} illustrates that Erasure almost never offers more useful information than any other mode.
Nonetheless, figure~\ref{fig:effort-table} does show that even if erasure often fails to provide helpful information, it provides the rational programmer with a quick answer.

Figure~\ref{fig:avo-matrix} also indicates that blame helps significantly over just using exception information.
Both Natural and Transient's blame modes improve over their corresponding exception modes by about ten percent.
This illustrates that blame tracking offers value for locating the source of mistakes beyond the value provided by the dynamic checks alone.
At only ten percent, however, the amount of added value suggests that foregoing blame tracking in order to obtain performance gains may not hurt debugging all too much.

Finally, the results show that while Natural clearly offers the most utility compared to the other modes, Transient is not too far behind.
Natural is more useful than every other mode in figure~\ref{fig:avo-matrix}, and by a larger proportion than any other mode.
That said, the same can be said of Transient's two blame modes as compared to every other mode but Natural.
Furthermore, the difference in usefulness between Natural and Transient's blame modes is quite small across the board.
As we point out in~\ref{sec:mutate}, however, we should note that our selection of scenarios for analysis introduces bias toward Erasure and Transient.

\subsection{Threats to validity}

@@

\subsection{Threat: Is our Transient Correct?}
\label{sec:threat:transient}
%% purpose: talk about transient-blame implementation .... what did we do,
%%  how did we test, what threats remain?
%%
%% the benchmark timeouts look like a threat, but this section shows that
%%  we've looked carefully at those

Surprisingly, Transient blames an empty set of boundaries in 967 cases.
Such cases should never occur, at least in theory, because an empty set
 means the value has never crossed a boundary---if the value is indeed defined
 in the current module, then we have typed code blaming itself for a typed
 value; thus, the type checker must be unsound.
After careful investigation of these empty blame cases, we found no soundness
 bugs in Typed Racket.
Instead, we found scenarios in which Transient lost track of the proper
 boundaries.
These scenarios suggest ways to improve the blame algorithm from~\citet{vss-popl-2017}:
\begin{itemize}
  \item
    Entries in a blame map must point to several parent entries.
    For example, if the function \texttt{f} receives bad input in the call
    \texttt{(filter f xs)}, then blame should point to both \texttt{filter}
    and the \texttt{xs} list.
  \item
    The construction of blame map entries must be guided by type-like specifications
     instead of relying on syntax.
    Aliasing the built-in \texttt{filter} function should not change the shape
     of the blame map.
  \item
    The initial blame map must reflect the initial type environment.
    Typed Racket trusts that untyped core-library functions, such as \texttt{filter},
     behave correctly, but does not list the assumption in the blame map.
\end{itemize}
\noindent{}Despite these known limitations, we conjecture that these improvements
 to Transient blame-tracking will not affect our overall conclusions.
